---
title: "Carnegie inundation"
author: "Adrian Pinder"
date: "22/11/2019"
output:
  html_document: 
    fig_caption: yes
    toc: yes
editor_options:
  chunk_output_type: console
---

RStudio: Version 1.0.136 2009-2016 RStudio Inc. R version: `r getRversion()`  
Date and time: `r Sys.time()`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_chunk$set(out.width='1.0\\linewidth')
```

```{r include=FALSE}
options(scipen=999)
```

```{r}
library(ggplot2)
library(gridExtra)
library(grid)
library(lattice)
library(scales)
```

####Data are remote sensed water presence, including variables inha (inundated hectares) and naha (not assessable hectares - where there are blank lines across the wetland polygon). naha areas are only added to the inha (to give a new column comha) where they are entirely across inha areas. Where they span both inha and dry areas (drha) they are not used. Code below creates files for each wetland polygon (site) for assessment (outside R) of which dates to use the naha data. Rainfall data is from Bureau of Meteorology.

Remote sensed inundated areas from from Bart Huntley (DBCA Remote Sensing and Spatial Analysis)

###Import inundation and rainfall data
```{r}
#read raw inundation data for all wetland polygons
Data <- read.csv("./Data/Carnegie inundation data.csv")
#Load daily rainfall data from BoM for Carnegie Station
Rainfall.Carn <- read.csv("./Data/Carnegie rainfall.csv")
Rainfall.Carn$Date <- as.Date(Rainfall.Carn$Date, "%d/%m/%Y")
#Load daily rainfall data from BoM for Prenti Station
Rainfall.Pren <- read.csv("./Data/Prenti rainfall.csv")
Rainfall.Pren$Date <- as.Date(Rainfall.Pren$Date, "%d/%m/%Y")
```

### Undertake correlations between rainfall stations
```{r}
#Spearman correlations
#omit dates with NA values for any station
Comb_Rain_3val <- na.omit(Comb_Rain)
#calculate spearman correlations
cor(Comb_Rain_3val$Carnegie.Rainfall, Comb_Rain_3val$Prenti.Rainfall, method = "spearman")
cor(Comb_Rain_3val$Prenti.Rainfall, Comb_Rain_3val$Wongawol.Rainfall, method = "spearman")
cor(Comb_Rain_3val$Carnegie.Rainfall, Comb_Rain_3val$Wongawol.Rainfall, method = "spearman")
```

### Calculations and Graphs for central lake area
```{r}
#extract main lake central data
Data_MLC <- Data[Data$site == "Main Lake Central", ]
#work out proportion of naha against inundated ha for examining 
Data_MLC$propnaha <- Data_MLC$naha/Data_MLC$inha*100
Data_MLC$date <- as.Date(Data_MLC$date, "%Y-%m-%d")
Data_MLC$nahause <- Data_MLC$naha
Data_MLC$naha_include <- "no"
#change naha_include to 'yes' for selected dates
Data_MLC$naha_include[Data_MLC$date %in% as.Date(c("07/01/2006","04/03/2009","10/03/2011","26/03/2011","11/04/2011","29/05/2011","14/06/2011","16/07/2011","01/08/2011","17/08/2011","25/08/2011","02/09/2011","18/09/2011","26/09/2011","05/11/2011","13/11/2011","15/03/2013","31/03/2013","16/04/2013","02/03/2014","19/04/2014"), "%d/%m/%Y")] <- "yes"
Data_MLC$nahause[Data_MLC$naha_include == "no"] <- 0
Data_MLC$comha <- Data_MLC$inha + Data_MLC$nahause
Data_MLC$propin <- Data_MLC$comha/32250.42*100
write.csv(Data_MLC, "Data_MLC.csv")
```

###Produce rainfall plots
```{r}
#combined Carnegie, Prenti and Wongawol rainfall - graphs and correlations
Comb_Rain <- read.csv("./Data/Combined rainfall.csv")
Comb_Rain$Date <- as.Date(Comb_Rain$Date, "%d/%m/%Y")
#create 3 ggplots - one for each rainfall station
RC <- ggplot () +
  geom_line(data = Comb_Rain, aes(x = Date, y = Carnegie.Rainfall), color = "blue", size = 0.5) + ylim(0,120) + theme(axis.title.x = element_blank())
RP <- ggplot () +
  geom_line(data = Comb_Rain, aes(x = Date, y = Prenti.Rainfall), color = "red", size = 0.5) + ylim(0,120) + theme(axis.title.x = element_blank())
RW <- ggplot () +
  geom_line(data = Comb_Rain, aes(x = Date, y = Wongawol.Rainfall), color = "green", size = 0.5) + ylim(0,120)
#arrange all three plots together
grid.arrange(RC, RP, RW, ncol=1)
```

##rainfall plots with central lake area rainfall for just 1999
```{r}
RC1999 <- ggplot () +
  geom_line(data = Comb_Rain, aes(x = Date, y = Carnegie.Rainfall), color = "blue", size = 0.5) + ylim(0,120) + theme(axis.title.x = element_blank()) + scale_x_date(date_breaks = "1 month", labels=date_format("%b"), limits = as.Date(c('1999-01-01','1999-12-01')))
RW1999 <- ggplot () +
  geom_line(data = Comb_Rain, aes(x = Date, y = Wongawol.Rainfall), color = "green", size = 0.5) + ylim(0,120) + scale_x_date(date_breaks = "1 month", labels=date_format("%b"), limits = as.Date(c('1999-01-01','1999-12-01')))
MLCarea1999 <- ggplot() +
  geom_line(data = Data_MLC, aes(x = date, y = propin), color = "red", size = 0.5) + ylim(0,120) + theme(axis.title.x = element_blank()) + scale_x_date(date_breaks = "1 month", labels=date_format("%b"), limits = as.Date(c('1999-01-01','1999-12-01')))
grid.arrange(RC1999, RW1999, MLCarea1999, ncol=1)
```



double line plot with rainfall x 150 so can see on same scale as inundated hectares
```{r}
ggplot () +
  geom_line(data = Data_MLC, aes(x = date, y = comha), color = "blue", size = 0.5) +
  geom_line(data = Rainfall.Carn, aes(x = Date, y = Rainfall*150), color = "red", size = 1, alpha=0.5) + scale_x_date(limits = as.Date(c('1999-01-01','1999-12-01')))
```

double line plot with rainfall x 150 so can see on same scale as inundated hectares (restrcited to 1999)
```{r}
ggplot () +
  geom_line(data = Data_MLC, aes(x = date, y = comha), color = "blue", size = 0.5) +
  geom_line(data = Rainfall.Carn, aes(x = Date, y = Rainfall*150), color = "red", size = 1, alpha=0.5) + scale_x_date(limits = as.Date(c('1999-01-01','1999-12-01')))
```

double line plot with different axes using par and base R plot.
```{r}
par(mar = c(5,5,2,5))
with(Rainfall.Carn, plot(Date, Rainfall, type="l", col = "dark green", axes=F, xlab=NA, ylab=NA, cex=0.8))
axis(side = 4)
mtext(side = 4, line = 3, 'Daily rainfall (mm)')

par(new = T)
with(Data_MLC, plot(date, comha, type="l", col= rgb(1,0,0, alpha=0.5), lwd=2,
             ylab=expression("inudated hectares")),
             ylim=c(0,32000))
```

### Calculations and Graphs for west lake area
```{r}
#extract main lake west data
Data_MLW <- Data[Data$site == "Main Lake West", ]
#work out proportion of naha against inundated ha for examining 
Data_MLW$propnaha <- Data_MLW$naha/Data_MLW$inha*100
Data_MLW$date <- as.Date(Data_MLW$date, "%Y-%m-%d")
Data_MLW$nahause <- Data_MLW$naha
Data_MLW$naha_include <- "no"
write.csv(Data_MLW, "Data_MLW.csv")
#change naha_include to 'yes' for selected dates
Data_MLW$naha_include[Data_MLW$date %in% as.Date(c("15/03/2013","11/04/2011","26/03/2011","10/03/2011","04/03/2009"), "%d/%m/%Y")] <- "yes"
Data_MLW$nahause[Data_MLW$naha_include == "no"] <- 0
Data_MLW$comha <- Data_MLW$inha + Data_MLW$nahause
Data_MLW$propin <- Data_MLW$comha/63194.49*100
```

double line plot with rainfall x 150 so can see on same scale as inundated hectares
```{r}
ggplot () +
  geom_line(data = Data_MLW, aes(x = date, y = comha), color = "blue", size = 0.5) +
  geom_line(data = Rainfall.Carn, aes(x = Date, y = Rainfall*150), color = "red", size = 1, alpha=0.5)
```

double line plot with different axes using par and base R plot.
```{r}
par(mar = c(5,5,2,5))
with(Rainfall.Carn, plot(Date, Rainfall, type="l", col = "dark green", axes=F, xlab=NA, ylab=NA, cex=0.8))
axis(side = 4)
mtext(side = 4, line = 3, 'Daily rainfall (mm)')

par(new = T)
with(Data_MLW, plot(date, comha, type="l", col= rgb(1,0,0, alpha=0.5), lwd=2,
             ylab=expression("inudated hectares")),
             ylim=c(0,32000))
```

### Calculations and Graphs for east lake area
```{r}
#extract main lake east data
Data_MLE <- Data[Data$site == "Main Lake East", ]
#work out proportion of naha against inundated ha for examining 
Data_MLE$propnaha <- Data_MLE$naha/Data_MLE$inha*100
Data_MLE$date <- as.Date(Data_MLE$date, "%Y-%m-%d")
Data_MLE$nahause <- Data_MLE$naha
Data_MLE$naha_include <- "no"
#None of the naha data used because there is so much land area within the polygon
Data_MLE$nahause[Data_MLE$naha_include == "no"] <- 0
Data_MLE$comha <- Data_MLE$inha + Data_MLE$nahause
Data_MLE$propin <- Data_MLE$comha/35199.99*100
```

### Calculations and Graphs Peripheral wetland 1
```{r}
#extract data for peripheral wetland 1
Data_P1 <- Data[Data$site == "Peripheral 1", ]
#work out proportion of naha against inundated ha for examining 
Data_P1$propnaha <- Data_P1$naha/Data_P1$inha*100
Data_P1$date <- as.Date(Data_P1$date, "%Y-%m-%d")
Data_P1$nahause <- Data_P1$naha
Data_P1$naha_include <- "no"
#change naha_include to 'yes' for selected dates
Data_P1$naha_include[Data_P1$date %in% as.Date(c("17/08/2011","15/03/2013","26/03/2011","12/03/2012","11/04/2011","29/05/2011","4/03/2009","2/03/2014","9/02/2012","10/03/2011","1/08/2011","18/09/2011","19/01/2016","14/06/2011","16/07/2011","23/12/2011","2/09/2011","31/03/2013","5/11/2011","20/03/2009","25/02/2012"), "%d/%m/%Y")] <- "yes"
Data_P1$nahause[Data_P1$naha_include == "no"] <- 0
Data_P1$comha <- Data_P1$inha + Data_P1$naha
Data_P1$propin <- Data_P1$comha/732.33*100
```

### Calculations and Graphs Peripheral wetland 2
```{r}
#extract data for peripheral wetland 2
Data_P2 <- Data[Data$site == "Peripheral 2", ]
#work out proportion of naha against inundated ha for examining 
Data_P2$propnaha <- Data_P2$naha/Data_P2$inha*100
Data_P2$date <- as.Date(Data_P2$date, "%Y-%m-%d")
Data_P2$nahause <- Data_P2$naha
Data_P2$naha_include <- "no"
write.csv(Data_P2, "Data_P2.csv")
#change naha_include to 'yes' for selected dates
Data_P2$naha_include[Data_P2$date %in% as.Date(c("28/03/2006","13/04/2006","29/04/2006","15/05/2006","16/06/2006","2/07/2006","18/07/2006","3/08/2006","19/08/2006","20/09/2006","10/01/2007","10/03/2011","26/03/2011","11/04/2011","14/06/2011","1/08/2011","17/08/2011","18/09/2011","5/11/2011","7/12/2011","15/03/2013","31/03/2013","16/04/2013","19/06/2013","5/07/2013","6/08/2013","22/08/2013"), "%d/%m/%Y")] <- "yes"
Data_P2$nahause[Data_P2$naha_include == "no"] <- 0
Data_P2$comha <- Data_P2$inha + Data_P2$naha
Data_P2$propin <- Data_P2$comha/38.52*100
```






### Calculations and Graphs Peripheral wetland 3
```{r}
#extract data for peripheral wetland 3
Data_P3 <- Data[Data$site == "Peripheral 3", ]
#work out proportion of naha against inundated ha for examining 
Data_P3$propnaha <- Data_P3$naha/Data_P3$inha*100
Data_P3$date <- as.Date(Data_P3$date, "%Y-%m-%d")
Data_P3$nahause <- Data_P3$naha
Data_P3$naha_include <- "no"
#export data to manually decide which dates to use naha
write.csv(Data_P3, "Data_P3.csv")
#change naha_include to 'yes' for selected dates
Data_P3$naha_include[Data_P3$date %in% as.Date(c("09/05/2004","26/06/2004","13/08/2004","29/08/2004","30/09/2004","13/04/2006","29/04/2006","15/05/2006","16/06/2006","02/07/2006","18/07/2006","03/08/2006","19/08/2006","20/09/2006","06/10/2006","7/11/2006","23/11/2006","09/12/2006","25/12/2006","10/01/2007","28/11/2008","04/03/2009","20/03/2009","23/05/2009","08/06/2009","12/09/2009","10/03/2011","26/03/2011","11/04/2011","29/05/2011","14/06/2011","16/07/2011","01/08/2011","17/08/2011","02/09/2011","18/09/2011","05/11/2011","23/12/2011","09/02/2012","25/02/2012","12/03/2012","13/04/2012","29/04/2012","16/06/2012","02/07/2012","03/08/2012","19/08/2012","04/09/2012","20/09/2012","07/11/2012","23/11/2012","10/01/2013","26/01/2013","11/02/2013","15/03/2013","31/03/2013","16/04/2013","19/06/2013","05/07/2013","06/08/2013","22/08/2013","23/09/2013","09/10/2013","29/01/2014","02/03/2014","19/04/2014","06/06/2014","24/07/2014","09/08/2014","25/08/2014","26/09/2014","12/10/2014","29/11/2014","31/12/2014","19/01/2016"), "%d/%m/%Y")] <- "yes"
Data_P3$nahause[Data_P3$naha_include == "no"] <- 0
Data_P3$comha <- Data_P3$inha + Data_P3$naha
Data_P3$propin <- Data_P3$comha/148.95*100
```

double line plot for peripheral wetland 3 with different axes using par and base R plot.
```{r}
par(mar = c(5,5,2,5))
with(Data_P3, plot(date, comha, type="l", col="red3", lwd=2,
             ylab=expression("inudated hectares")),
             ylim=c(0,32000))

par(new = T)
with(Rainfall.Carn, plot(Date, Rainfall, type="l", col = rgb(0,0.6,0.2, alpha=0.5), axes=F, xlab=NA, ylab=NA, cex=1.2))
axis(side = 4)
mtext(side = 4, line = 3, 'Daily rainfall (mm)')
```

### Calculations and Graphs Peripheral wetland 4
```{r} 
#extract Peripheral wetland 4 data
#since no occasions on which we would add in the naha data, the original inha data is used
Data_P4 <- Data[Data$site == "Peripheral 4", ]
Data_P4$date <- as.Date(Data_P4$date, "%Y-%m-%d")
Data_P4$propin <- Data_P4$inha/214.38*100
```

### Calculations and Graphs Peripheral wetland 5
```{r}
#extract data for peripheral wetland 5
Data_P5 <- Data[Data$site == "Peripheral 5", ]
#work out proportion of naha against inundated ha for examining 
Data_P5$propnaha <- Data_P5$naha/Data_P5$inha*100
Data_P5$date <- as.Date(Data_P5$date, "%Y-%m-%d")
Data_P5$nahause <- Data_P5$naha
Data_P5$naha_include <- "no"
write.csv(Data_P5, "Data_P5.csv")
#change naha_include to 'yes' for selected dates
Data_P5$naha_include[Data_P5$date %in% as.Date(c("26/07/2003","27/08/2003","26/06/2004","13/08/2004","13/04/2006","10/03/2011","26/03/2011","11/04/2011","29/05/2011","14/06/2011","16/07/2011","1/08/2011","17/08/2011","2/09/2011","18/09/2011","15/03/2013","31/03/2013","16/04/2013","12/12/2013","2/03/2014","19/04/2014","21/05/2014"), "%d/%m/%Y")] <- "yes"
Data_P5$nahause[Data_P5$naha_include == "no"] <- 0
Data_P5$comha <- Data_P5$inha + Data_P5$naha
Data_P5$propin <- Data_P5$comha/328.5*100
```

double line plot with different axes using par and base R plot.
```{r}
par(mar = c(5,5,2,5))
with(Data_P4, plot(date, inha, type="l", col="red3", lwd=2,
             ylab=expression("inudated hectares")),
             ylim=c(0,200))

par(new = T)
with(Rainfall.Carn, plot(Date, Rainfall, type="l", col = rgb(0,0.6,0.2, alpha=0.5), axes=F, xlab=NA, ylab=NA, cex=1.2))
axis(side = 4)
mtext(side = 4, line = 3, 'Daily rainfall (mm)')
```

#create grid of three inundation plots
```{r, fig.width=10, fig.height=20}

#create 8 ggplots - one for each wetland polygon
MLC <- ggplot () +
  geom_line(data = Data_MLC, aes(x = date, y = propin), color = "blue", size = 0.5) + theme(axis.title.x = element_blank()) + ylab("% Inundated")  + theme(axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0), size=9)) + labs(title = "Central area of main lake") + theme(plot.title = element_text(size = 9)) + scale_y_continuous(limits = c(0,100),  breaks = c(50, 100))
MLW <- ggplot () +
  geom_line(data = Data_MLW, aes(x = date, y = propin), color = "blue", size = 0.5) + theme(axis.title.x = element_blank()) + ylab("% Inundated")  + theme(axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0), size=9)) + labs(title = "Western area of main lake") + theme(plot.title = element_text(size = 9)) + scale_y_continuous(limits = c(0,100),  breaks = c(50, 100))
MLE <- ggplot () +
  geom_line(data = Data_MLE, aes(x = date, y = propin), color = "blue", size = 0.5) + theme(axis.title.x = element_blank()) + ylab("% Inundated")  + theme(axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0), size=9)) + labs(title = "Eastern area of main lake") + theme(plot.title = element_text(size = 9)) + scale_y_continuous(limits = c(0,100),  breaks = c(50, 100))
P1 <- ggplot () +
  geom_line(data = Data_P1, aes(x = date, y = propin), color = "orange", size = 0.5) + theme(axis.title.x = element_blank()) + ylab("% Inundated") + theme(axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0), size=9))+ labs(title = "Peripheral wetland 1") + theme(plot.title = element_text(size = 9)) + scale_y_continuous(limits = c(0,100),  breaks = c(50, 100))
P2 <-  ggplot () +
  geom_line(data = Data_P2, aes(x = date, y = propin), color = "black", size = 0.5) + theme(axis.title.x = element_blank()) + ylab("% Inundated") + theme(axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0), size=9))+ labs(title = "Peripheral wetland 2") + theme(plot.title = element_text(size = 9)) + scale_y_continuous(limits = c(0,100),  breaks = c(50, 100))
P3 <- ggplot () +
  geom_line(data = Data_P3, aes(x = date, y = propin), color = "red", size = 0.5) + theme(axis.title.x = element_blank()) + ylab("% Inundated") + theme(axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0), size=9))+ labs(title = "Peripheral wetland 3") + theme(plot.title = element_text(size = 9)) + scale_y_continuous(limits = c(0,100),  breaks = c(50, 100))
P4 <- ggplot () +
  geom_line(data = Data_P4, aes(x = date, y = propin), color = "green", size = 0.5) + theme(axis.title.x = element_blank()) + ylab("% Inundated") + theme(axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0), size=9)) + labs(title = "Peripheral wetland 4 on dune") + theme(plot.title = element_text(size = 9)) + scale_y_continuous(limits = c(0,100),  breaks = c(50, 100))
P5 <- ggplot () +
  geom_line(data = Data_P5, aes(x = date, y = propin), color = "purple", size = 0.5) + ylab("% Inundated") + theme(axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0), size=9)) + labs(title = "Peripheral wetland 5") + theme(plot.title = element_text(size = 9)) + scale_y_continuous(limits = c(0,100),  breaks = c(50, 100))
#arrange all three plots together
grid.arrange(MLC, MLW, MLE, P1, P2, P3, P4, P5, ncol=1)
```

```{r}
#create plot of % wetland inundation using grid arrange
jpeg("rplot.jpg", width = 350, height = 500)
# 2. Create the plot
#arrange all three plots together
grid.arrange(MLC, MLW, MLE, P1, P2, P3, P4, P5, ncol=1)
# 3. Close the file
dev.off()
```


