---
title: "Carnegie inundation"
author: "Adrian Pinder"
date: "22/11/2019"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

RStudio: Version 1.0.136 2009-2016 RStudio Inc. R version: `r getRversion()`  
Date and time: `r Sys.time()`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(out.width='1.0\\linewidth')
```

```{r include=FALSE}
options(scipen=999)
```

```{r}
library(ggplot2)
library(gridExtra)
library(grid)
library(lattice)
```

####Data are remote sensed water presence data including variables inha (inundated hectares) and naha (not assessable hectares - where there are blank lines across the wetland polygon). These naha areas are only added to the inha (to give a new column comha) where they are entirely across inha areas. Where they span both inha and dry areas they are not used. Code below creates files for each wetland polygon for assessment (outside R) of which dates to use the naha data.


```{r}
#read the raw inundation data for all wetland polygons from Bart Huntley (DBCA Remote Sensing and Spatial Analysis).
Data <- read.csv("Carnegie_inundation_data.csv")
#Load daily rainfall data from BoM for carnegie Station
Rainfall.Carn <- read.csv("Carnegie rainfall.csv")
Rainfall.Carn$Date <- as.Date(Rainfall.Carn$Date, "%d/%m/%Y")
#Load daily rainfall data from BoM for Prenti Station
Rainfall.Pren <- read.csv("Prenti rainfall.csv")
Rainfall.Pren$Date <- as.Date(Rainfall.Pren$Date, "%d/%m/%Y")
```

```{r}
#combined Carnegie, Prenti and Wongawol rainfall - graphs and correlations
Comb_Rain <- read.csv("Combined rainfall.csv")
Comb_Rain$Date <- as.Date(Comb_Rain$Date, "%d/%m/%Y")

#create 3 ggplots - one for each rainfall station
RC <- ggplot () +
  geom_line(data = Comb_Rain, aes(x = Date, y = Carnegie.Rainfall), color = "blue", size = 0.5) + ylim(0,120) + theme(axis.title.x = element_blank())
RP <- ggplot () +
  geom_line(data = Comb_Rain, aes(x = Date, y = Prenti.Rainfall), color = "red", size = 0.5) + ylim(0,120) + theme(axis.title.x = element_blank())
RW <- ggplot () +
  geom_line(data = Comb_Rain, aes(x = Date, y = Wongawol.Rainfall), color = "green", size = 0.5) + ylim(0,120)
#arrange all three plots together
grid.arrange(RC, RP, RW, ncol=1)

#omit dates with NA values for any station
Comb_Rain_3val <- na.omit(Comb_Rain)
#calculate spearman correlations
cor(Comb_Rain_3val$Carnegie.Rainfall, Comb_Rain_3val$Prenti.Rainfall, method = "spearman")
cor(Comb_Rain_3val$Prenti.Rainfall, Comb_Rain_3val$Wongawol.Rainfall, method = "spearman")
cor(Comb_Rain_3val$Carnegie.Rainfall, Comb_Rain_3val$Wongawol.Rainfall, method = "spearman")
```

```{r}
par(mar = c(5,5,2,5))
with(Data_P4, plot(date, inha, type="l", col="red3", lwd=2,
             ylab=expression("inudated hectares")),
             ylim=c(0,200))

par(new = T)
with(Rainfall.Carn, plot(Date, Rainfall, type="l", col = rgb(0,0.6,0.2, alpha=0.5), axes=F, xlab=NA, ylab=NA, cex=1.2))
axis(side = 4)
mtext(side = 4, line = 3, 'Daily rainfall (mm)')





## Calculations and Graphs for central lake area
```{r}
#extract main lake central data
Data_MLC <- Data[Data$site == "Main Lake Central", ]
#work out proportion of naha against inundated ha
Data_MLC$propnaha <- Data_MLC$naha/Data_MLC$inha*100
#create copy for editing in excel - save edited version as Data_MLC2 - edits are whether to use the naha data (include_naha = yes or no)
write.csv(Data_MLC, "Data_MLC.csv")
#read back in data
Data_MLC2 <- read.csv("Data_MLC2.csv")
```

```{r}
#create new variable for the naha data to use by copying the naha data
Data_MLC2$nahause <- Data_MLC2$naha
#if include_naha = no then set nahause to 0
Data_MLC2$nahause[Data_MLC2$include_naha == "no"] <- 0
#sum nahause and inha to give combined inundated area
Data_MLC2$comha <- Data_MLC2$inha + Data_MLC2$nahause
Data_MLC2$date <- as.Date(Data_MLC2$date, "%d/%m/%Y")
```

double line plot with rainfall x 150 so can see on same scale as inundated hectares
```{r}
ggplot () +
  geom_line(data = Data_MLC2, aes(x = date, y = comha), color = "blue", size = 2) +
  geom_line(data = Rainfall.Carn, aes(x = Date, y = Rainfall*150), color = "red", size = 1, alpha=0.5)
```

double line plot with different axes using par and base R plot.
```{r}
par(mar = c(5,5,2,5))
with(Rainfall.Carn, plot(Date, Rainfall, type="l", col = "dark green", axes=F, xlab=NA, ylab=NA, cex=0.8))
axis(side = 4)
mtext(side = 4, line = 3, 'Daily rainfall (mm)')

par(new = T)
with(Data_MLC2, plot(date, comha, type="l", col= rgb(1,0,0, alpha=0.5), lwd=2,
             ylab=expression("inudated hectares")),
             ylim=c(0,32000))
```

## Calculations and Graphs Peripheral wetland 3
```{r}
#extract main lake central data
Data_P3 <- Data[Data$site == "Peripheral 3", ]
#work out proportion of naha against inundated ha
Data_P3$propnaha <- Data_P3$naha/Data_P3$inha*100
#create copy for editing in excel - save edited version as Data_MLC2 - edits are whether to use the naha data (include_naha = yes or no)
write.csv(Data_P3, "Data_P3.csv")
#read back in data
Data_P3.2 <- read.csv("Data_P3_2.csv")
```

```{r}
#create new variable for the naha datato use by copying the naha data
Data_P3.2$nahause <- Data_P3.2$naha
#if include_naha = no then set nahause to 0
Data_P3.2$nahause[Data_P3.2$include_naha == "no"] <- 0
#sum nahause and inha to give combined inundated area
Data_P3.2$comha <- Data_P3.2$inha + Data_P3.2$nahause
Data_P3.2$date <- as.Date(Data_P3.2$date, "%d/%m/%Y")
```

double line plot with different axes using par and base R plot.
```{r}
par(mar = c(5,5,2,5))
with(Data_P3.2, plot(date, comha, type="l", col="red3", lwd=2,
             ylab=expression("inudated hectares")),
             ylim=c(0,32000))

par(new = T)
with(Rainfall.Pren, plot(Date, Rainfall, type="l", col = rgb(0,0.6,0.2, alpha=0.5), axes=F, xlab=NA, ylab=NA, cex=1.2))
axis(side = 4)
mtext(side = 4, line = 3, 'Daily rainfall (mm)')
```

## Calculations and Graphs Peripheral wetland 4
```{r} 
#extract Peripheral wetland 4 data
#since no occasions on which we would add in the naha data, the original inha data is used
Data_P4 <- Data[Data$site == "Peripheral 4", ]
Data_P4$date <- as.Date(Data_P4$date, "%Y-%m-%d")
```

double line plot with different axes using par and base R plot.
```{r}
par(mar = c(5,5,2,5))
with(Data_P4, plot(date, inha, type="l", col="red3", lwd=2,
             ylab=expression("inudated hectares")),
             ylim=c(0,200))

par(new = T)
with(Rainfall.Carn, plot(Date, Rainfall, type="l", col = rgb(0,0.6,0.2, alpha=0.5), axes=F, xlab=NA, ylab=NA, cex=1.2))
axis(side = 4)
mtext(side = 4, line = 3, 'Daily rainfall (mm)')
```


